A full approximation scheme multilevel method \\ for solving nonlinear variational inequalities

### 2024-07-21 talk-ismp/slides.tex
practice this version?; need to get below 25 min

	\begin{itemize}
	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $F : \mathcal{V} \to \mathcal{V}'$ continuous, $\ell\in \mathcal{V}'$
	\item[$\circ$] example is the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
	\end{itemize}
\item discretization gives algebraic system on fine mesh $\Omega^h$:
    $${\color{FireBrick} F^h(u^h) = \ell^h}$$
\end{itemize}

@@

\begin{itemize}
\item FAS multigrid works \alert{very well} on nice nonlinear PDE problems
\item example: Liouville-Bratu equation
    $$-\nabla^2 u - e^u = 0$$
with Dirichlet boundary conditions on $\Omega=(0,1)^2$
\item minimal problem-specific code:
    \begin{enumerate}
    \item[1.] residual evaluation on grid level: $F^j(\cdot)$
    \item[2.] pointwise smoother: $\phi_i(c) = 0 \,\forall i$
        \begin{itemize}
        \item[$\circ$] nonlinear Gauss-Seidel iteration
        \end{itemize}
    \item[3.] coarsest-level solve can be same as smoother
    \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{the meanings of ``fast solver''}

\begin{itemize}
\item what does ``very well'' on the previous slide mean?
\begin{block}{definition} a solver is \emph{optimal} if work is $O(N)$ for $N$ unknowns
\end{block}
    \begin{itemize}
    \item[$\circ$] measure work in flops or run-time
    \item[$\circ$] constant iterations of $O(N)$ method $\implies$ optimal
    \end{itemize}
\end{itemize}

\begin{itemize}
\item multigrid people get greedy:
\begin{block}{definition} a solver shows \emph{textbook multigrid efficiency} if it does total work less than 10 times that of a single smoother sweep
\end{block}
    \begin{itemize}
    \item[$\circ$] TME $\implies$ optimal
    \end{itemize}
\end{itemize}
\end{frame}

@@

\subsection{results}  % FIXME lead with results

@@

\begin{itemize}
\item new algorithm (Bueler \& Farrell 2023):

@@

\begin{itemize}
\item start with subspace decomposition over nested meshes:
    $$\Omega^j \subset \Omega^{j+1}$$
\item the FE function spaces $\mathcal{V}^j$ over $\Omega^j$ are also nested:
    $$\mathcal{V}^j \subset \mathcal{V}^{j+1}$$

\begin{block}{definition}
$\ds \mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ \quad is called a \emph{subspace decomposition} (Xu 1992)
\end{block}
    \begin{itemize}
    \item[$\circ$] \emph{non}-unique vector space sum
    \item[$\circ$] Xu's paper explains how to analyze linear multigrid for PDEs via subspace decomposition
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Tai's (2003) constraint decomposition \emph{non-trivially} extends a subspace decomposition $\mathcal{V}^J = \sum_i\mathcal{V}^i$ to convex subsets
\item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
\begin{block}{definition}
$\ds \mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$ \quad is a \emph{constraint decomposition} (CD) if there are closed and convex subsets $\mathcal{K}^i\subset \mathcal{V}^i$, and (nonlinear) projections $\Pi_i : \mathcal{K}^J \to \mathcal{K}^i$, so that $\ds v = \sum_{i=0}^J \Pi_i v$ and a stability condition applies (not shown)
\end{block}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD $\mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$

@@

\medskip
\item Tai's iterations are not practical because you compute on the finest level in fact
\item we added two techniques: \emph{defect obstacles} on each level, and \emph{FAS coarse corrections}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item recall $\mathcal{K} = \{v \ge \psi\}$ in an obstacle problem
\begin{block}{definition}
for finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$ and an iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} (Gr\"aser \& Kornhuber 2009) is
    $$\chi^J = \psi^J - w^J \in \mathcal{V}^J$$
\end{block}

    \begin{itemize}
    \item[$\circ$] note $\chi^J \le 0$
    \end{itemize}
\item we generate the CD through

@@

    \begin{itemize}
    \item[$\circ$] a \emph{nonlinear} operator
    \item[$\circ$] {\footnotesize also due to (Gr\"aser \& Kornhuber 2009)}
    \end{itemize}
\end{itemize}

@@

\begin{frame}{up and down sets}

\begin{itemize}
\item upward part in the FASCD V-cycle uses large admissible sets:
    $$\mathcal{U}^j = \{z^j \ge \chi^j\}$$
\item downward sets are smaller to guarantee admissibility of the upcoming coarse correction:
    $$\mathcal{D}^j = \{y^j \ge \phi^j=\chi^j - \chi^{j-1}\}$$
\item $\ds \mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ is a CD of the $j$th-level admissible set
\end{itemize}

\begin{center}
\includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
\end{center}
\end{frame}

\begin{frame}{multilevel constraint decomposition in FASCD}

\centering
\includegraphics[width=0.7\textwidth]{../talk-dms/figs/innerconeapprox.png}
\end{frame}

\begin{frame}{full approximation scheme constraint decomposition}

% case of lower obstacle only
\begin{pseudo}[font=\small]
\pr{fascd-vcycle}(J,\ell^J,\psi^J;w^J)\text{:} \\+
    $\chi^J = \psi^J - w^J$ \\
    for $j=J$ downto $j=1$ \\+
      $\chi^{j-1} = \maxR \chi^j$ \\
      $\phi^j = \chi^j - P\chi^{j-1}$ \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$ \\
      $w^{j-1} = \iR(w^j + y^j)$ \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$  \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$  \\-
    return $w^J+z^J$
\end{pseudo}
\end{frame}

\begin{frame}{FASCD V-cycle: visualization on a 1D problem}

@@

\begin{frame}{FASCD specifics}

see paper (Bueler \& Farrell 2023) for:
\begin{itemize}
\item generalization to upper and lower obstacles:
    $$\mathcal{K}^J = \{\underline{\psi}^J \le v^J \le \overline{\psi}^J\}$$
\item stopping criteria
    \begin{itemize}
    \item[$\circ$] evaluate whether CP/KKT conditions are satisfied
    \end{itemize}
\item FMG cycle
\item details of $O(m_J)$ smoother
\end{itemize}
\end{frame}

% FIXME move earlier \section{results}

\only<1>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/bothitersVF.png}} \hfill
\only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
\only<4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
\end{frame}

\begin{itemize}
\item suppose $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $0\le u\le 1$
\item suppose it moves by combination of diffusion, advection by wind $\bm{X}(x)$, and source function $\phi(x)$:
    $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$$
\item two active sets (\emph{$d=2$ case}):
    $$\underline{A}_u = \{u(x) = 0\} \hspace{22mm} \overline{A}_u = \{u(x) = 1\} \hspace{8mm}$$
\end{itemize}

\begin{center}
\includegraphics[width=0.75\textwidth]{../talk-dms/figs/advdiff.png}
\end{center}

\scriptsize
\begin{itemize}
\item[] \emph{compare: linear programming (Klee-Minty cube?), spatial correlations}
\end{itemize}
\end{frame}

\begin{itemize}
\item ``where are there glaciers?'' is a free-boundary problem
\end{itemize}

\vspace{-2mm}
\hfill {\tiny Sequinot et al.~(2018)}
\end{frame}

\begin{itemize}
\item glacier = incompressible, viscous fluid driven by gravity
\item to find: ice surface elevation $s(t,x,y)$ and velocity $\bu(t,x,y,z)$
\item over fixed bed topography with elevation $b(x,y)$
    \begin{itemize}
    \item[$\circ$] $s(t,x,y) \ge b(x,y)$
    \end{itemize}
\item in a \emph{climate} which adds or removes ice at a signed rate $a(t,x,y)$
    \begin{itemize}
    \item[$\circ$] data $a,b$ is defined on domain $\Omega \subset \RR^2$
    \end{itemize}
\end{itemize}

\begin{frame}{glacier free-boundary problem: naive strong form}

\begin{itemize}
\item is this an adequate description?:
\begin{align*}
s &\ge b                    & &\text{everywhere in } \Omega \\
\frac{\partial s}{\partial t} &= a + \bu|_s \cdot \bn_s & &\text{where } s(t,x,y) > b(x,y)
\end{align*}
\item notes:
    \begin{itemize}
    \item[$\circ$] surface velocity $\bu|_s$ is, \alert{in some manner}, determined by $s$
        \begin{itemize}
        \item $\bu|_s$ is generally a \emph{non-local} function of $s$
        \end{itemize}
    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
    \end{itemize}
\end{itemize}

\bigskip
\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
\end{frame}

\begin{itemize}
\item admissible surface elevations:\footnote{$(s-b)^{8/3} \stackrel{?}{\in} W^{1,4}(\Omega)$ so $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \dots  see (Jouvet \& Bueler, 2012)}
    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
\item steady ($\frac{\partial s}{\partial t}=0$) VI problem for surface elevation $s\in\mathcal{K}$:
	$$\ip{\Phi(s) - a}{r-s} \ge 0 \quad \text{ for all } r \in \mathcal{K}$$
where
    $$\Phi(s)=- \bu|_s \cdot \bn_s$$
with extension by 0 to all of $\Omega$
%\item $\bu$ is the velocity solution on
%    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\} \subset \RR^3$$
\end{itemize}
\end{frame}

\begin{frame}{shallow ice approximation}

\begin{itemize}
\item the \emph{shallow ice approximation} is a highly-simplified view of conservation of momentum
\item isothermal, nonsliding case:
\begin{align*}
\Phi(s) &= - \bu|_s \cdot \bn_s \\
        &= - \frac{\gamma}{4} (s-b)^{4} |\grad s|^{4} - \grad \cdot\left(\frac{\gamma}{5} (s-b)^{5} |\grad s|^{2} \grad s\right)
\end{align*}
\end{itemize}
\end{frame}

\begin{frame}{FASCD test case: simplified ice sheet the size of Greenland}

\begin{itemize}
\item \emph{ice sheet} = big glacier
\end{itemize}

\bigskip\bigskip
\centering
\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
\end{frame}

\begin{itemize}
\item observed optimality of FMG solver
\item good parallel \emph{weak scaling} as well
    \begin{itemize}
    \item[$\circ$] each processor owns $641\times 641$ (sub) mesh
    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns
    \item[] \dots and 88 meter resolution
    \end{itemize}
\end{itemize}

\bigskip
\centering
\includegraphics[width=0.55\textwidth]{../talk-dms/figs/siaweaktime.png}
\end{frame}

@@

\begin{itemize}
\item FASCD = new multilevel solver for VI (free-boundary) problems
    \begin{itemize}
    \item[$\circ$] implemented in Python Firedrake (over PETSc)
    \end{itemize}
\item observed optimality, even TME, in many cases
    \begin{itemize}
    \item[$\circ$] actually fast
    \end{itemize}

\bigskip\bigskip
\item[] \textbf{{\color{FireBrick} to do:}}
    \begin{itemize}
    \item add mesh adaptivity to free boundary (Stefano)
    \item implement in C inside PETSc
    \item apply to space-time (\emph{parabolic}) VI problems
    \item prove convergence
    \item identify smoothers for problems like elastic contact
    \item include membrane stresses in glacier case
    \end{itemize}
\end{itemize}
\end{frame}

>>>>

	\begin{itemize}
	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $F : \mathcal{V} \to \mathcal{V}'$ continuous, $\ell\in \mathcal{V}'$
	\item[$\circ$] example is the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
	\end{itemize}
\item discretization\footnote{discretization = finite element (FE) method, throughout this talk} gives algebraic system on fine mesh $\Omega^h$:
    $${\color{FireBrick} F^h(u^h) = \ell^h}$$
\end{itemize}

@@

\begin{itemize}
\item FAS multigrid is a fast solver on nice nonlinear PDE problems
\item example: Liouville-Bratu equation
    $$-\nabla^2 u - e^u = 0$$
with Dirichlet boundary conditions on $\Omega=(0,1)^2$
\item minimal problem-specific code:
    \begin{enumerate}
    \item[1.] residual evaluation on grid level: $F^j(\cdot)$
    \item[2.] pointwise smoother: $\phi_i(c) = 0 \,\forall i$
        \begin{itemize}
        \item[$\circ$] nonlinear Gauss-Seidel iteration
        \end{itemize}
    \item[3.] coarsest-level solve can be same as smoother
    \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{a meaning for ``fast solver''}

\begin{block}{definition} a solver is \emph{optimal} if work is $O(N)$ for $N$ unknowns
\end{block}

\bigskip
\begin{itemize}
\item one can measure work in flops or run-time
\item constant iterations of $O(N)$ method $\implies$ optimal
\end{itemize}
\end{frame}

@@

\subsection{results, and an application}

\only<1>{recall:

\centerline{\includegraphics[width=0.9\textwidth]{../talk-oxford/images/obstacle65.pdf}}}
\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<5>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill
\only<2-4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
\only<5>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}

\only<2-5>{\noindent \scriptsize
\emph{all results from Bueler \& Farrell (2024)}}
\end{frame}

\begin{itemize}
\item $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $\boxed{0\le u\le 1}$
\item substance moves by diffusion, advection (wind $\bm{X}(x)$), and source ($\sigma(x)$):
    $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \sigma$$
\item two active sets:
    $$\underline{A}_u = \{u(x) = 0\} \hspace{22mm} \overline{A}_u = \{u(x) = 1\} \hspace{8mm}$$
\end{itemize}

\begin{center}
\includegraphics[width=0.75\textwidth]{figs/advdiff.png}
\end{center}

\vspace{-3mm}
\hfill \tiny \emph{Bueler \& Farrell (2024)}

\medskip
\scriptsize
\begin{itemize}
\item<2> compare linear programming (Klee-Minty cube?) \dots with strong spatial correlations
\end{itemize}
\end{frame}

\begin{itemize}
\item ``where are the glaciers?'' is a free-boundary problem
\end{itemize}

\vspace{-2mm}
\hfill {\tiny Seguinot et al.~(2018)}
\end{frame}

\begin{itemize}
\item glacier = incompressible, viscous fluid driven by gravity
\item in a \emph{climate} which adds or removes ice at a signed rate $a(t,x,y)$
\item over fixed bed topography with elevation $b(x,y)$
    \begin{itemize}
    \item[$\circ$] data $a,b$ is defined on domain $\Omega \subset \RR^2$
    \end{itemize}
\item to find: ice surface elevation $s(t,x,y)$ and velocity $\bu(t,x,y,z)$
\item subject to constraint $\boxed{s\ge b}$
\end{itemize}

\begin{itemize}
\item admissible surface elevations:
    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
\item steady-state VI problem for surface elevation $s\in\mathcal{K}$:
	$$\ip{\Phi(s) - a}{r-s} \ge 0 \quad \text{ for all } r \in \mathcal{K}$$
where $\Phi(s)=- \bu|_s \cdot \bn_s$
\item in the isothermal, nonsliding, \emph{shallow ice approximation}\footnote{$=$ \emph{a highly-simplified view of conservation of momentum, in which} $(s-b)^{8/3} \in W^{1,4}(\Omega)$ \emph{so} $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \hfill \emph{(Jouvet \& Bueler, 2012)}} case:
\begin{align*}
\Phi(s) &= - \bu|_s \cdot \bn_s \\
        &= - \frac{\gamma}{4} (s-b)^{4} |\grad s|^{4} - \grad \cdot\left(\frac{\gamma}{5} (s-b)^{5} |\grad s|^{2} \grad s\right)
\end{align*}
\end{itemize}
\end{frame}

\begin{frame}{test case: ice sheet ($=$ big glacier) the size of Greenland}

\centering
\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
\end{frame}

\begin{itemize}
\item optimality of FMG solver (not shown)
\item good parallel \emph{weak scaling}
    \begin{itemize}
    \item[$\circ$] each processor owns $641\times 641$ (sub) mesh
    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns
    \item[] \dots and 88 meter resolution
    \end{itemize}
\end{itemize}

\bigskip
\centering
\includegraphics[width=0.55\textwidth]{../talk-dms/figs/siaweaktime.png}

\hfill \scriptsize
\emph{Bueler \& Farrell (2024)}
\end{frame}

@@

\begin{itemize}
\item new algorithm (Bueler \& Farrell 2024):

@@

\begin{itemize}
\item start with subspace decomposition over nested meshes:
    $$\Omega^j \subset \Omega^{j+1}$$
\item the FE vector spaces $\mathcal{V}^j$ over $\Omega^j$ are also nested:
    $$\mathcal{V}^j \subset \mathcal{V}^{j+1}$$
\item $\mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ is called a \emph{subspace decomposition} (Xu 1992)

    \begin{itemize}
    \item[$\circ$] \emph{non}-unique vector space sum
    \item[$\circ$] Xu's paper explains how to analyze linear multigrid for PDEs via subspace decomposition
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Tai's (2003) constraint decomposition \emph{non-trivially} extends subspace decomposition to convex subsets
\item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
\begin{block}{definition}
$\ds \mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$ \quad is a \emph{constraint decomposition} (CD) if there are closed and convex subsets $\mathcal{K}^i\subset \mathcal{V}^i$, and (nonlinear) projections $\Pi_i : \mathcal{K}^J \to \mathcal{K}^i$, so that $\ds v = \sum_{i=0}^J \Pi_i v$ and a stability condition applies (not shown)
\end{block}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD $\mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$, specifically:

@@

\item Tai's iterations are not practical because you must evaluate all residuals on the finest level
\item two techniques make a practical solver:

    \begin{itemize}
    \item[$\circ$] \emph{defect obstacles} on each level  \hfill {\scriptsize (Gr\"aser \& Kornhuber (2009)}
    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize (Bueler \& Farrell (2024)}
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item suppose $\mathcal{K} = \{v \ge \psi\}$ in an obstacle problem
\begin{block}{definition}
for finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$ and an iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} is
    $$\chi^J = \psi^J - w^J \in \mathcal{V}^J$$
\end{block}

    \begin{itemize}
    \item[$\circ$] note $\chi^J \le 0$
    \end{itemize}
\item generate the CD through

@@

    \begin{itemize}
    \item[$\circ$] a \emph{nonlinear} operator
    \end{itemize}
\end{itemize}

@@

\begin{frame}{up and down CDs in FASCD}

\begin{center}
\includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
\end{center}

\begin{itemize}
\item upward direction in V-cycle uses larger admissible sets:
    $$\mathcal{U}^j = \{z^j \ge \chi^j\}$$
\item downward sets are smaller to guarantee admissibility of the upcoming coarse correction:
    $$\mathcal{D}^j = \{y^j \ge \phi^j=\chi^j - \chi^{j-1}\}$$
\item $\ds \mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ is a CD of the $j$th-level admissible set
\end{itemize}
\end{frame}

\begin{frame}{up and down CDs in FASCD}

\begin{center}
\includegraphics[width=0.7\textwidth]{../talk-dms/figs/innerconeapprox.png}
\end{center}

\begin{itemize}
\item all sets $\mathcal{U}^j$ and $\mathcal{D}^j$ are cones which include the origin
\end{itemize}
\end{frame}

\begin{frame}{FASCD = full approximation scheme constraint decomposition}

\vspace{-2mm}
\begin{pseudo}[font=\small]
\pr{fascd-vcycle}(J,\ell^J,\psi^J;w^J)\text{:} \\+
    $\chi^J = \psi^J - w^J$ \\
    for $j=J$ downto $j=1$ \\+
      $\chi^{j-1} = \maxR \chi^j$ \\
      $\phi^j = \chi^j - P\chi^{j-1}$ \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$ \\
      $w^{j-1} = \iR(w^j + y^j)$ \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$  \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$  \\-
    return $w^J+z^J$
\end{pseudo}

\vspace{-2mm}
\tiny
\begin{itemize}
\item[] \emph{(unilateral (lower obstacle only) version shown for simplicity)}
\end{itemize}
\end{frame}

\begin{frame}{V-cycle: visualization on a 1D problem}

@@

\begin{frame}{specifics}

see (Bueler \& Farrell 2024) for:
\begin{itemize}
\item generalization to upper and lower obstacles:
    $$\mathcal{K}^J = \{\underline{\psi}^J \le v^J \le \overline{\psi}^J\}$$
\item stopping criteria
    \begin{itemize}
    \item[$\circ$] evaluates CP/KKT conditions
    \end{itemize}
\item FMG cycle
\item details of $O(m_J)$ smoother
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item FASCD = new multilevel solver for VI (free-boundary) problems
    \begin{itemize}
    \item[$\circ$] implemented in Python Firedrake (over PETSc)
    \end{itemize}
\item observed optimality in many cases
    \begin{itemize}
    \item[$\circ$] actually fast on tested nonlinear PDE problems
    \end{itemize}

\bigskip\bigskip
\item many things \textbf{{\color{FireBrick} to do}}:
    \begin{itemize}
    \item[$\circ$] implement in C inside PETSc
    \item[$\circ$] prove convergence; compare (Reusken, 1988)
    \item[$\circ$] apply to space-time (\emph{parabolic}) VI problems
    \item[$\circ$] add mesh adaptivity
    \item[$\circ$] identify smoothers for problems like elastic contact
    \item[$\circ$] include membrane stresses in glacier case
    \end{itemize}
\end{itemize}
\end{frame}
